Matching bullets

### 2015-11-21 imaging-paper.Rnw
biasotti added to references

subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
subLOFx1$y <- subLOFx1$y + 19*1.5625 # working now!!!
lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
@
\begin{figure}
\centering
    \subfloat[Crosscuts at $x = 100$.\label{fig:crosscut}]{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer("", palette="Set1") + theme_bw() +
  theme(legend.position = c(.9, .2))
@
    }
    \hfill
    \subfloat[Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}]{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer(palette="Set1") + theme_bw() +
  theme(legend.position="none")
@
    }
\caption{\label{fig:cross100} Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}
\centering
    \subfloat[Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}]{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.49\\textwidth', warning=FALSE>>=
lofX <- lofX %>% group_by(bullet) %>% mutate(
  l30 = smoothloess(y, resid, span = 0.03)
)
lofX$r05 <- 0.75* sign(lofX$l30) * as.numeric(abs(lofX$l30) > .75)
lofX$type <- factor(lofX$r05)
levels(lofX$type) <- c("groove", NA, "peak")

qplot(data=lofX, y, resid, colour=bullet, size=I(0.5), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey30", size=0.75) +
  facet_grid(bullet~.) +
  scale_colour_brewer("", palette="Set1") +
  theme_bw() +
  theme(legend.position="none")
@
    }
    \hfill
    \subfloat[Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}]{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='.49\\textwidth', warning=FALSE>>=
qplot(data=lofX, x=y, y=l30, colour=type, geom="line", 
      group=bullet, size = I(0.75)) + 
  facet_grid(bullet~.) +
  ylab("Individual characteristics") + 
#  geom_point(aes(y=r05)) +
  geom_rect(xmin = 200, xmax = 2260, ymin = -.75, ymax = 0.75, colour=NA, fill=alpha("grey90", alpha =0.5)) +
  scale_colour_brewer(palette="Set2", na.value = "grey90") +
  theme_bw() +
  theme(legend.position="bottom")
@
    }
\hfill
    \subfloat[Overlaps of the same color indicate matching striation marks.  \label{fig:smoothcutc}]{%
<<smoothtile, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=2.5, out.width='.49\\textwidth'>>=
qplot(data=lofX, x=y, y=bullet, fill=type, geom="tile") +
  scale_fill_brewer(palette="Set2", na.value="grey90") + 
  theme_bw() +
  ylab("") +
  theme(legend.position="bottom")
@
    }    
\caption{\label{fig:match} Matching striation marks: smooth (a), cut (b), and match (c).}
\end{figure}

\hh{Figure~\ref{fig:match} gives an overview of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; XXX need to vary over this threshold level next). Two lines are defined as a match, if their peak/groove base at the threshold overlaps. If there is no overlap, a peak/groove leads to a mismatch. We then count the number of maximum consecutive matches. }

>>>>

subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))

# smooth
lofX <- lofX %>% group_by(bullet) %>% mutate(
  l30 = smoothloess(y, resid, span = 0.03)
)
# cut at .75
lofX$r05 <- 0.75* sign(lofX$l30) * as.numeric(abs(lofX$l30) > .75)
lofX$type <- factor(lofX$r05)
levels(lofX$type) <- c("groove", NA, "peak")
@

<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
matches <- lofX %>% group_by(y) %>% summarise(
  potential = (length(unique(type)) == 1),
  allnas = sum(is.na(type))/n(),
  type1 = na.omit(type)[1],
  type = paste(type, sep="|", collapse="|"),
  n = n()
)

matches$id <- cumsum(matches$allnas == 1) + 1
matches$lineid <- as.numeric(matches$allnas != 1) * matches$id

isMatch <- function(id, type) {
  if (id[1] == 0) return(FALSE)
#  browser()
  types <- strsplit(type, split = "|", fixed=TRUE) 
  t1 <- sapply(types, function(x) x[1])
  t2 <- sapply(types, function(x) x[2])
  if (all(t1 == "NA")) return(FALSE)
  if (all(is.na(t2))) return(FALSE)
  if (all(t2 == "NA")) return(FALSE)

  peak <- length(grep("peak", c(t1, t2))) > 0
  groove <- length(grep("groove", c(t1, t2))) > 0
  if (peak & groove) return(FALSE)

  return(TRUE)
}

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)

@
\begin{figure}
\centering
    \subfloat[Crosscuts at $x = 100$.\label{fig:crosscut}]{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer("", palette="Set1") + theme_bw() +
  theme(legend.position = c(.9, .2))
@
    }
    \hfill
    \subfloat[Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}]{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer(palette="Set1") + theme_bw() +
  theme(legend.position="none")
@
    }
\caption{\label{fig:cross100} Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}
\centering
    \subfloat[Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}]{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.49\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=bullet, size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey30", size=0.5) +
  facet_grid(bullet~.) +
  scale_colour_brewer("", palette="Set1") +
  theme_bw() +
  theme(legend.position="none")
@
    }
    \hfill
    \subfloat[Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}]{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='.49\\textwidth', warning=FALSE>>=
qplot(data=lofX, x=y, y=l30, colour=type, geom="line", 
      group=bullet, size = I(0.75)) + 
  facet_grid(bullet~.) +
  ylab("Individual characteristics") + 
#  geom_point(aes(y=r05)) +
  geom_rect(xmin = 200, xmax = 2260, ymin = -.75, ymax = 0.75, colour=NA, fill=alpha("grey90", alpha =0.5)) +
  scale_colour_brewer(palette="Set2", na.value = NA) +
  theme_bw() +
  theme(legend.position="bottom")
@
    }
\hfill
    \subfloat[Grey backgrounds identify a line on one of the bullets. Overlaps of the same color indicate matching striation marks.
    Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}]{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.49\\textwidth'>>=
ggplot() +
  geom_rect(aes(xmin = miny, xmax = maxy), ymin = 0.25, ymax=2.75, fill="grey80", data = subset(lines, lineid != 0)) +
  geom_tile(data=lofX, aes(x=y, y=bullet, fill=type)) + 
  scale_y_discrete(expand = c(0.2,0)) +
  scale_fill_brewer(palette="Set2") + 
  theme_bw() +
  ylab("") +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
  geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
@
    }    
\caption{\label{fig:match} Matching striation marks: smooth (a), cut (b), and match (c).}
\end{figure}

\hh{Figure~\ref{fig:match} gives an overview of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; XXX need to vary over this threshold level next). Two lines are defined as a match, if their peak/groove base at the threshold overlaps. If there is no overlap, a peak/groove leads to a mismatch. We then count the number of maximum consecutive matches. In this example, the number of consecutive matching striations (CMS) is fifteen, a very high number indicative of a match between the bullets. For lead bullets, such as this, Biasotti \cite{biasotti:1959} XXX need more references XXX considered four or more consecutive lines to be sufficient evidence of a match. }

\hh{XXX next step: how does this matching work in known non-matches? distribution of CMS under a null of no match? }
